The Galactic Midplane Is Not a Plane: Implications for Dynamical Analysis with {\em Gaia} Data and Beyond

### 2019-04-26 paper/ms.tex
Small changes throughout

The second issue, which is related to the point of this paper, is that based
on the results in our previous sections computing $\bm{J}_i$ for a field star
and directly comparing it to  $\bm{J}_c$ will have significant systematic
error if the cluster and field star are far enough apart to have significantly
different local midplanes (see Figures \ref{fig:midplane} and \ref{fig:midplane_chervin}). Since our criterion for cluster membership (distance in action space) scales with the
cluster mass, the region where the search will succeed varies with cluster
mass as well. To illustrate this effect, we assume a cluster mass $m_c$ and
estimate the expected action-space dispersion from
Equation~\eqref{eq:action_disp}. We then consult
Figure~\ref{fig:many_orbit_wrong_ref} to determine the midplane offset
necessary in order for the induced error in the vertical action to be the same
as the expected action-space dispersion. We perform this analysis for the
thin-disk-like and thick-disk-like orbits considered in
Section~\ref{ssec:quant}, assuming that the halo-like orbit is unlikely for
members of known open clusters (see Table~\ref{tab:real_clusters}). The result
for the allowable midplane offset for these two orbits in shown in
Figure~\ref{fig:cluster_offset} as a function of cluster mass.

@@

\item To demonstrate the previous point, we measured the local galactic
midplane along the Solar circle in three different high-resolution, zoom-in
simulations of Milky Way mass galaxies from the FIRE collaboration. We found
that the midplane varies as a function of azimuth by $\sim185, 162, 84\,\pc$ (middle $90\%$) for the
three galaxies we considered (\mi{}, \mf{}, and \mm{}, respectively).

@@

\acknowledgments
We would like to thank Megan Bedell, Tobias Buck, Elena D'Onghia, Benoit
Famaey, and Adrian Price-Whelan for helpful discussions. A.B. would like to
thank Todd Phillips for helpful discussions. A.B. was supported in part by the
Roy \& Diana Vagelos Program in the Molecular Life Sciences and the Roy \&
Diana Vagelos Challenge Award. M.-M.M.L. was partly supported by NSF grant
AST18-15461. The work of D.A.-A., A.B., D.W.H., C.F.P.L., R.S., K.V.J.,
M.-M.M.L., and M.K.N. is supported by the Simons Foundation.

>>>>

The second issue, which is related to the point of this paper, is that based
on the results in our previous sections computing $\bm{J}_i$ for a field star
and directly comparing it to  $\bm{J}_c$ will have significant systematic
error if the cluster and field star are far enough apart to have significantly
different local midplanes (see Figures~\ref{fig:midplane}~and~\ref{fig:midplane_chervin}). Since our criterion for cluster membership (distance in action space) scales with the
cluster mass, the region where the search will succeed varies with cluster
mass as well. To illustrate this effect, we assume a cluster mass $m_c$ and
estimate the expected action-space dispersion from
Equation~\eqref{eq:action_disp}. We then consult
Figure~\ref{fig:many_orbit_wrong_ref} to determine the midplane offset
necessary in order for the induced error in the vertical action to be the same
as the expected action-space dispersion. We perform this analysis for the
thin-disk-like and thick-disk-like orbits considered in
Section~\ref{ssec:quant}, assuming that the halo-like orbit is unlikely for
members of known open clusters (see Table~\ref{tab:real_clusters}). The result
for the allowable midplane offset for these two orbits in shown in
Figure~\ref{fig:cluster_offset} as a function of cluster mass.

@@

\item To demonstrate the previous point, we measured the local galactic
midplane along the Solar circle in three different high-resolution, zoom-in
simulations of Milky Way mass galaxies from the FIRE collaboration. We found
that the midplane varies as a function of azimuth at the Solar circle by $\sim185, 162, 84\,\pc$ (middle $90\%$) for the
three galaxies we considered (\mi{}, \mf{}, and \mm{}, respectively).

@@

\acknowledgments
We would like to thank Megan Bedell, Robert A. Benjamin, Tobias Buck, Elena D'Onghia, Benoit
Famaey, and Adrian Price-Whelan for helpful discussions. A.B. would like to
thank Todd Phillips for helpful discussions. A.B. was supported in part by the
Roy \& Diana Vagelos Program in the Molecular Life Sciences and the Roy \&
Diana Vagelos Challenge Award. M.-M.M.L. was partly supported by NSF grant
AST18-15461. The work of D.A.-A., A.B., D.W.H., C.F.P.L., R.S., K.V.J.,
M.-M.M.L., and M.K.N. is supported by the Simons Foundation.
